Measurement of a dwarf galaxy rotational velocity from its Lyman-alpha line emission

### 2016-08-09 paper/letter.tex
streamlined physical interpretation

In the rotation model the size of that region can be infered from the
constraint on the total HI mass.
From the optical depth of $10^7$ and the non-detection in the HI line,
we have an upper limit for the size where the Lya emission comes of
$D<0.34$kpc.

 For an homogeneous sphere the HI optical depth from its can be
 written as $\tau = \sigma_0 n D/2$, where $\sigma_0=5.898\times
10^{-14}$cm$^{-2}$ is the Lyman$\alpha$  optical depth at the 
line's center, $n$ is the number density and $D$ is the sphere's
diameter. 
From this we can impose additional constrains on $D$ from the typical
values of the Hydrogen number density and our constrain on
$\tau=10^{7}$.  Using a range of $1<n/\mathrm{atoms/cm}^{-3} <
10^{-3}$. This gives us a range of $0.11 < D/\mathrm{kpc}<100$. 
Together from the total HI mass we have thus that the HI region should
have a diameter of $0.11 < D/\mathrm{kpc}<0.34$.

This can be rewritten in terms of the gas' temperature $T$ and column
density $N_{H}$as $\tau = 3.31 \times 10^{-14} (10^{4}\mathrm{K}/T)^{1/2}
(N_{H}/\mathrm{atoms\ cm}^{-2})$.

This allows us to approximate the total hydrogen mass as
\begin{equation}
M_{H} = m_{H}  N_{H} D^{2} = 226\times \tau  \left(\frac{T}{10^4
  \mathrm{K}}\right)^{1/2}\left(\frac{D}{\mathrm{kpc}}\right)^2M_{\odot}
\end{equation}

On the other hand, we have an estimate for the dynamical mass from the
galaxy size $D$ and its rotational velocity $V$:

\begin{equation}
M_{T} = \frac{V^{2}D}{G} = 2.16\times10^{5}
\left(\frac{V}{\mathrm{km\ s}^{-1}}\right)^2\left(\frac{D}{\mathrm{kpc}}\right) M_{\odot}
\end{equation}

From this limit and the rotational velocity of $300$ km/s and the
limit in the size $D$ we have limits of for the dynamical mass of
$2.1\times 10^{9}$M$_{\odot}$ $<M_D<$  $6.6\times 10^{9}$M$_{\odot}$.
which is at least $7$ to $25$ times larger than the H$I$ mass. The
mass to luminosity ratio is in turn between $2$ to $7$ times
$L_{\odot}/M_{\odot}$ considering the total luminosity in the $V$
band.

>>>>

In the rotation model the size of the spherical region, $r_{s}$, can
be infered from the relationship $\tau = \sigma_0 n r_s$, where
$\sigma_0=5.898\times 10^{-14}$cm$^{-2}$ is the Lyman$\alpha$
cross section at the  line's center, $n$ is the number density. 
This expresion can be rewritten as
\begin{equation}
r_{s} = 0.055 \left(\frac{\tau}{10^7}\right) \left(
\frac{\mathrm{atoms\ cm^{-3}}}{n}\right) \mathrm{kpc}.
\end{equation}
%
Using this result and assuming a bound for the number density of
$10^{-2}<n/\mathrm{atoms\ cm^{-3}}<1$ we have $5.5\times 10^{-2}
< r_s/\mathrm{kpc}<5.5$.

This is consistent with the  constraint on the total HI mass and the
optical depth $\tau_0$ as we show next. The total HI mass can be
approximated as $M_H = \frac{4\pi}{3} m_H n r_s^3$, where
$m_H=1.67\times 10^{-24}$ gr is the mass of a single Hydrogen atom.
This allows us to write the total hydrogen mass as
\begin{equation}
M_{H} = 5.70 \times 10^{6} \left(\frac{\tau}{10^{7}}\right)
\left(\frac{r_s}{\mathrm{kpc}}\right)^2 M_{\odot}
\end{equation}
%
The uppper limit from the HI mass non-detection imposes $r_{s} <
6.81$ kpc, which is consistent with the bound we found in the first
place, as we wanted to show.

Using those size bounds for the spherical model we find a constrain for
the dynamical mass of $5.45\times 10^{8}M_{\odot} < M_{\rm dyn} <
5.45\times 10^{10} M_{\odot}$.
